### 2021-09-15 paper.Rmd
some progress

---
title: |
  | Emotional framing and the effectiveness of 
  | corrective information^[I thank Thomas Bräuninger, Sebastian Juhl, the GIP team, and participants of the Fall 2020 CDSS workshop in political science at the University of Mannheim for excellent comments and helpful suggestions on this study. This work was supported by the German Research Foundation (DFG) via the Collaborative Research Center (SFB) 884 on ``The Political Economy of Reforms'' (Project C2) and the University of Mannheim’s Graduate School of Economics and Social Sciences.]
  | *Working Paper*
author: |
  | *Working paper*
  |
  | Pirmin Stöckle^[Corresponding address: pirmin.stoeckle@gess.uni-mannheim.de]
  | GESS and SFB 884, University of Mannheim
date: \today
linestretch: 1.5
colorlinks: true
abstract: \setstretch{1} \noindent Concerns about various forms of misinformation and its fast dissemination through online media have generated huge interest into ways to effectively correct false claims. An under-explored mechanism in this research is the role of distinct emotions. How do emotional appeals interact with corrective information? Specifically, I focus on the emotion of disgust, which has been shown to be linked to the moralization of attitudes, which in turn reduces the impact of empirical evidence on attitudes and makes compromise less likely. Substantively, I investigate the issue of genetically modified (GM) food. I implement a pre-registered survey experiment within a panel study based on a probability sample of the general population in Germany (N = 4,000). Results show what we have learned about the question. This study provides evidence on the effect of emotionally charged disinformation on perceptions of GM food, and ways to effectively correct false claims. In a broader perspective, these results inform further studies and policy interventions on other issues where disinformation loads on strong emotions, ranging from social policy over immigration to health interventions such as vaccinations. \vspace{.8cm}
# abstract-template: Everyone agrees that this issue is really important. But we do not know much about this specific question, although it matters a great deal, for these reasons. We approach the problem from this perspective. Our research design focuses on these cases and relies on these data, which we analyze using this method. Results show what we have learned about the question. They have these broader implications.

bibliography: references.bib # to be downloaded from citavi or google drive
csl: american-journal-of-political-science.csl # https://github.com/citation-style-language/styles
output:
  bookdown::pdf_document2:
    toc: no
    keep_tex: true
mainfont: Times New Roman
sansfont: Times New Roman
fontsize: 12pt
link-citations: true
documentclass: article
geometry: margin=1in
always_allow_html: yes

---

<!-- This paper is written in RMarkdown based on a template available here: https://osf.io/q395s -->

```{r, include=FALSE}
# packages
library(tidyverse)
library(broom)
library(scales)
library(modelsummary)
library(gridExtra)
library(jtools)
library(tidyverse)
library(sjmisc)
library(scales)
library(knitr)
```

\newpage
# Introduction
<!-- Guiding Question about people, not scholars! -->
How do emotional appeals interact with corrective information?
<!-- context on misinformation and the important role of emotions in spreading it -->
Fueled by the fast dissemination through digital media, the issue of various forms of misinformation has become a major concern both in academia and for public policy [@jerit2020political; @lazer2018science; @lewandowsky2012misinformation; @wardle2017information].
To limit the danger from inaccurate or misleading information, efforts of fact-checking or corrective information more generally are widely applied [@walter2020fact]. A common finding in the maturing literature on corrective information is that these interventions are usually successful in reducing misperceptions, but may not translate to other outcomes such as attitudes, candidate evaluations or blame attribution [@bisgaard2019getting; @nyhan2019taking; @swire2019they; @thorson2016belief].

<!-- % still, we know little about the role of discrete emotions in interacting with misinformation and corrective information -->
While this has been linked to affect [@thorson2016belief], the connection of distinct emotions to both susceptibility to misinformation, and the effectiveness of corrective information is under-explored (exceptions include [@ecker2011terrorists; @martel2020reliance; @rosenzweig2021happiness; @weeks2015emotions]. This is unfortunate given the important role that emotional appeals play in the dissemination of false or misleading information. We know that false news spread faster than true news, possibly because they are more likely to evoke emotions of surprise and disgust [@vosoughi2018spread]. In general, moral-emotional language is rewarded on social media [@brady2017emotion] and populist parties use more negative emotional appeals in their communication than non-populist parties [@widmann2021how].

<!-- % emotion in political science -->
In this study, I follow recent work in political science inspired by moral psychology. Most relevant for this study, @clifford2019emotional finds that a persuasive emotional frame loading on emotions of anger and disgust increases the perception of an issue as being one of moral conviction [@skitka2020psychology; @wisneski2017moralization]. 
Moral conviction, i.e. the conviction that the handling of an issue touches morality and is about fundamental right or wrong, can be seen as a dimension of attitude strength [@skitka2018attitude].

<!-- % get more concrete linking belief echoes / affective perseverance and specific emotion with possible connection to moralization -->
In his study, Clifford shows that disgust-evoking messages can increase moral conviction with its consequences on polarizing attitudes. However, some of the treatment material used in his study comes from a later retracted study linking GMO and cancer. Such a piece of information can be classified as misinformation in the sense of going against ``best available evidence in the public sphere''[@flynn2017nature, p.128].
From findings of belief echoes [@thorson2016belief] and affective perseverance [@sherman2002affective] we might suspect that such emotional effects of frames may persist even in the face of successful factual correction.

<!-- % Contribution bringing belief echoes and discrete emotions together -->
This study contributes to the literature of corrective information by linking the concept of belief echoes to the perseverance of specific emotions with related consequences on attitudes. Specifically, it provides evidence on the effect of emotionally charged disinformation on perceptions of GM food, and ways to effectively correct false claims.

<!-- % Final Punchline on broader relevance -->
In the view of many scientists [@roberts2018nobel, planck2019scientists], genetic engineering provides avenues with large potential benefits, which may be impeded by public resistance possibly originating from misleading claims easily disseminated through online media.
In a broader perspective, these results inform further studies and policy interventions on other issues where disinformation loads on strong emotions, ranging from social policy over immigration to health interventions such as vaccinations.

# Theory

## Misinformation and corrective information {-}

<!-- % some definitions from Kuklinski and Flynn to Jerit -->

<!-- % information corrects factual beliefs, but the effects of information on attitudes are mixed -->

Factual beliefs are affected in the direction of presented evidence for both (a) misinformation and (b) corrective information.

The effects on related opinions are much less clear.
Some studies find effects, some not.

Affective responses are at least implicitly discussed in this literature, but discrete emotions are widely neglected.

## Emotions, political learning, moralization {-}

<!-- % From Vasiloupolou / Wagner -->
<!-- %Early research into emotional reactions either used one simple positive–negative dimension or a two-dimensional valence model (Marcus, 2003; Marcus et al., 2000). More recently, researchers have argued that distinct emotions may characterise individuals’ reactions (e.g. Huddy et al., 2007; Larsen and McGraw, 2011; Petersen, 2010; Smith et al., 2008; Valentino et al., 2008, 2011; Weber, 2012; see also Conover and Feldman, 1986). Appraisal theories of emotions suggest that various distinct emotions arise due to individuals’ cognitive interpretation of the situation they experience (e.g. Roseman, 1991; Smith and Ellsworth, 1985; for a review, see Brader and Marcus, 2013). Neuroscientific theories based on the neural processes that generate emotional responses have also increasingly turned towards a three-dimensional model of emotions that distinguishes between fear, anger and enthusiasm (Brader and Marcus, 2013; Lerner and Keltner, 2001). Research in political psychology building on both appraisal and neural process theories has therefore focused in particular on the causes and consequences of anger, fear1 and enthusiasm as distinct emotions (Carver and Harmon-Jones, 2009; Conover and Feldman, 1986; Huddy et al., 2005, 2007; Lerner and Keltner, 2000, 2001; Petersen, 2010; Smith and Ellsworth, 1985; Valentino et al., 2011). -->
<!-- % -->
<!-- %Applied to political phenomena, distinct emotions cause different types of actions and attitudes: fear leads to caution and openness, and anger to more confident and aggressive responses (Druckman and McDermott, 2008; MacKuen et al., 2010). For instance, whereas fear and anger both tend to increase intolerance and threat perceptions (e.g. Erişen and Kentmen-Cin, 2017), they lead to distinct ways of reacting to terrorist threats (Lerner et al., 2003). Fearful citizens are more wary of the increased risk and turn to isolationism, while angry citizens are less risk-averse and tend to favour pro-active intervention (Huddy et al., 2007). Anxious citizens are also more likely to seek out new information and process this carefully (Brader et al., 2008; Marcus et al., 2000; Merolla and Zechmeister, 2009; Valentino et al., 2008). Finally, enthusiasm increases the willingness to participate in politics, but like anger it also increases less careful forms of information-processing (Brader, 2006; Marcus et al., 2000). -->

Emotions certainly play an important role in politics [@redlawsk2006feeling], and approaches such as cognitive appraisal theory [@moors2013appraisal] or affective intelligence theory [@marcus2000affective; @marcus2019applying] provide frameworks for empirical expectaitoins about specific emotions.

Although there is a rich literature on emotions in politics shedding light on its effect on information processing, political learning and participation, the connection of distinct emotions to both susceptibility to misinformation and the effectiveness of corrective information is under-explored.

Research has shown that... anger... anxiety... enthusiasms...

A recently growing interest is in disgust.
Connection to morality, relevant for immigration, homeless ...

A study has shown that moral shock works to moralize attitudes towards GM food.

While it is often seen as a binary categorization whether or not an issue is a moral one, there is quite some variation in the extent of moralization within issues [@ryan2014reconsidering] with distinct behavioral consequences, such as an absolute unwillingness to compromise [@ryan2017no] or a tendency to view that issue in absolutist term and to be less likely to be open for information about consequences [@ryan2019actions].

However, the main part of the material used there can be classified as misinformation
GM food is an issue on which disinformation exists. This does not mean that there can or should be no criticism, but concrete adverse health effects are not warranted by evidence. Least knowledgeable people are the fiercest opponents. So

## Empirical expectations {-}

From findings on moral conviction and how emotional frames induce it [@clifford2019emotional; @wisneski2017moralization], I expect that a persuasive message evoking negative emotions, notably disgust and anger, can increase perceptions of the respective issue as being one of moral conviction. In order to count as a persuasive message, I expect the message to affect both factual beliefs and policy opinions in its direction.

* H1 A persuasive, emotionally framed message evoking negative emotions of disgust and anger makes individuals more likely to favor policy restrictions on the respective issue and to perceive the issue as one of moral conviction.

<!-- % Moderation -->

<!-- % possible theoretical contribution by categorizing different emotion effects -->
<!-- %To categorize effects of emotions within the context mis- and corrective information, there are two general ways in which emotions may affect its effectiveness. In both cases, one can further differentiate between cases where the emotion is invoked by, or at least closely tied to, the object that is part of the information (integral emotion), and cases where the emotion is induced irrespective of the piece of information (incidental emotion, often also referred to as mood). Emotion as moderator. emotions affect information processing so also for misinformation and corrective information. But anxiety, anger and disgust word in different directions... emotion within fake news, see Martel. Emotion as outcome -->

Concerning the impact of corrective information, I expect that it will be effective in reducing factual misperceptions, irrespective of the presence of evoked emotions. This is in line with previous research showing no moderation effect on corrections due to emotional framing  [@ecker2011terrorists].

* H2a Corrective information is effective in eliminating the misinformation effect on factual beliefs.

<!-- % Belief echoes and affective perseverance -->

However, similar to previous results on affective perseverance [@sherman2002affective] and belief echoes [@thorson2016belief], I expect the effects of emotionally framed misinformation on policy opinion, moral conviction, and emotions to persist even in the face of effective factual correction.

* H2b The emotionally framed misinformation leaves an impact on policy opinions, moral conviction, and emotions.

Together, hypotheses 2a and 2b represent the existence of a ``belief echo'', i.e. the continued effect of a piece of misinformation on attitudinal or emotional outcomes, despite successful factual correction. If only H2a is supported, the correction is altogether successful.

<!-- % Counter-framing -->

Corrective information such as fact-checking is usually deliberately presented in a rather neutral way. In the case of strongly emotional framed misinformation, such a neutral correction may thus be successful in correcting factual beliefs, but not be able to counter the emotional impact of the misinformation. If this is the case, a more effective strategy may be to counter the emotional misinformation not with a neutral correction but with a correction embedded in an emotional counter-frame. If equally persuasive, such counterframes have been shown to negate the effect of an original frame [@chong2013counterframing]. I expect a similar effect to hold in response to emotionally framed misinformation.

* H3 A correction embedded in a persuasive emotional counter-frame eliminates the misinformation effect on policy opinions, moral conviction, and emotions.

<!-- %Visual cues relate to \citep{hameleers2020picture} and \citep{powell2021selection} -->

## GMO - a case for emotional reactions and misinformation {-}

<!-- % explain the relevance of GMO as an issue, also in realtion to moralization -->
The topic of GM feed is particularly suited for this study because it is an issue where science and public opinion diverge considerably, implying a relevant role for corrective information [@diamond2020does, hasell2020differential, carnahan2019processing]. This is especially true within the German population, which is where this study draws its sample from [@gaskell2010europeans]. GM food is also a topic on which there are misperceptions, possibly due to disinformation. Of course, there is legitimate criticism on the use of genetic modification. However, studies show that extreme opponents of GM food know the least about it [@fernbach2019extreme], and a lack of acceptance of novel food technologies against scientific evidence can be seen as as a threat for the development of a more resilient food system [@siegrist2020consumer].

# Data & Methods

To investigate the direct effects of misinformation and corrective information conditional on emotional framing, I conducted a pre-registered survey experiment.^[Pre-registration can be found on OSF: https://osf.io/y63xu] The experiment was implemented within wave 53 of the German Internet Panel (GIP) [@blom2015setting]. The GIP is a longitudinal panel survey based on a probability sample of the general population in Germany. The survey is conducted online, but recruitment was conducted offline according to a strict random probability sample. Households that did not posses the technical devices to participate online have been equipped with such devices. Fieldwork for this study took place in May 2021 and the sample size reached 4,260 respondents.

The between-subjects survey experiment followed a 3 $\times$ 3 full factorial design varying both misinformation (none, low emotional framing, high emotional framing) and corrective information (none, neutral correction, counterframe correction). Respondents were randomly assigned to one of the 9 resulting groups, thus yielding 8 treatment conditions and one control condition were no information treatment was shown.

An extensive debriefing is included at the end of the study for all respondents. Specifically, all original sources of information are revealed.
Because all shown information has in a similar format been in the media, there is no deception taking place.
The experimental treatment intends to evoke emotions, but I consider them within reason. There are two main reasons for that. First, the material including the used imagery has been used in a previously published study @[clifford2019emotional]. Second, the central part of the visual material has been used by actual reporting on the issue in mainstream media^[www.welt.de/newsticker/news1/article109594074/Bundesinstitut-Genmais-Schaeden-nicht-ausreichend-belegt].
The study was approved by the team operating the GIP, with the goal of designing the study in a way that it is an appropriately pleasant experience for respondents. Because this approval has been granted, this study did not require further approval by an ethics committee.

## Treatment Material {-}

The experimental treatment consists of a combination of two types of information. One type of information reports the results of a study claiming that genetically modified corn lead to increased cancer rates in rats. The wording is partly taken from actual news coverage of the study in 2012, and partly from treatment material in @clifford2019emotional. The respective study has been retracted due to non-adherence to scientific procedures, and the current scientific evidence does not point to severe health problems associated with genetically modified animal feed. The information within the study can thus be classified as misinformation.^[This is consistent with a definition of resulting mis*perceptions* as "factual beliefs that are false or contradict the best available evidence in the public domain" [@flynn2017nature, p.128]. If one would presume the intent to mislead, this would then be classified as disinformation.] I vary this misinformation to be either low or high in emotional framing by ex- or including graphical images of a rat with a cancer tumor and rotten corn, images that are taken from @clifford2019emotional and shown to induce negative feelings, specifically disgust.

The second type of information reports the accurate scientific consensus that GM food is generally safe to eat, that a study showing cancer growth in rats due to GM feed has been retracted, and that genetically modifying plants for the food industry is not a new phenomenon. This piece of information is a correction to the misinformation in that it reflects best available scientific evidence and it is constructed to conform to best practice standards of debunking information @[lewandowsky2020debunking]. As a neutral correction, this piece of information corrects the misinformation about concrete health risks from GM food in a text-only format. By negating the risks, it still stays within the overall framing of GM food as possibly risky, and does not add specific further information about benefits from GM food. This is done in a further condition, where the corrective information is embedded in a counter-frame, emphasizing positive aspects of GM food, such as its potential to contribute to more sustainable agriculture by reducing pesticides, and its potential to fight malnutrition.

Assignment is done by simple random assignment into equal sized groups. Table gives an overview over the nine resulting experimental groups. The misinformation (if any is assigned) is always shown first, immediately followed by the corrective information (if any is assigned). For respondents, the manipulation is thus designed in a way that both types of information ``hits them between the eyes'' @[kuklinski2000misinformation, p.805].

## Outcome measures {-}

As the factual belief of interest, I measure factual beliefs about the safety of GM foods by asking respondents whether they think that the statement that eating GM foods is safe for one's health is false or correct on a 7-point scale from "certainly false" to "certainly correct".

To provide a broad measure of policy support, I use a measure by @[diamond2020does] that asks respondents how high a policy priority it should be to restrict the use of GM food on a 7-point scale from "no priority" to "highest priority".

I measure moral conviction with two standard items from the literature, asking respondents whether their evaluation of GM food is connected to moral convictions and about fundamental right and wrong, respectively, on a 5-point scale from "not at all" to "very much" @[skitka2020psychology]. This is a direct measure of a "meta-cognition", building on the assumption that respondents know which issues are about moral conviction for them. Both items will be combined into a simple additive index.

I measure distinct emotions by directly asking respondents how well a given feeling represents how they feel toward GM food. Asking about their feelings toward GM food makes this a measure of integral, not incidental emotion, in line with expectation about to which kind of emotion moralization responds @[skitka2018attitude]. The distinct negative emotions I measure are disgust, anger, anxiety, and sadness, each on a 5-point scale from "not at all" to "extremely".

# Results

```{r, warning=FALSE, echo=FALSE}
# load data
source("data-preparation.R")
```

## Descriptive statistics {-}

Table \@ref(tab:descriptives)

```{r descriptives, echo=FALSE, message=FALSE, warning=FALSE}
library(modelsummary)
# only control group

data %>% 
  filter(exp_group == "control") %>% 
  select("Anger" = "anger",
         "Anxiety" = "anxiety",
         "Disgust" = "disgust",
         "Sadness" = "sadness",
         "Belief GMO safety" = "knowledge_safe", 
         "Policy restrictions" = "policy_opinion",
         "Moral conviction" = "moral_conviction") %>% 
  datasummary_skim(histogram =FALSE,
                   output = "default",
                   title = "Descriptive statistics in the control group")

```

## Moralization through misinformation {-}

To quantify the effects of the different treatment conditions on the outcomes of interest, I run separate linear regressions with the respective outcome as the dependent variable and the treatment group indicator as the independent variable.

```{r, echo = FALSE}
# linear regression models with 0-1 rescaled DV
lm_knowledge_safe <- lm(rescale(knowledge_safe) ~ exp_group, data)
lm_policy_opinion <- lm(rescale(policy_opinion) ~ exp_group, data)
lm_moral_conviction <- lm(rescale(moral_conviction) ~ exp_group, data)

lm_emotion_disgust <- lm(rescale(disgust) ~ exp_group, data)
lm_emotion_anger <- lm(rescale(anger) ~ exp_group, data)
lm_emotion_anxiety <- lm(rescale(anxiety) ~ exp_group, data)
lm_emotion_sadness <- lm(rescale(sadness) ~ exp_group, data)

# define functions
# function to get coefficients for plotting
get_pdata <- function(model, positions = c(), label = NA) {
  df <- data.frame(summary(model)$coefficients[positions,],
                   group = str_remove(rownames(summary(model)$coefficients[positions,]), "exp_group"),
                   outcome = label)
  df$group <- fct_inorder(df$group)
  df$outcome <- fct_inorder(df$outcome)
  names(df) <- c("estimate", "se", "t", "p", "group", "outcome")
  df$lower <- df$estimate - 1.96*df$se
  df$upper <- df$estimate + 1.96*df$se
  row.names(df) <- NULL
  return(df)
}

```

```{r fig-1, fig.align="center", fig.cap="Coefficients (and 95 confidence intervals) from OLS-regressions with the outcome variable rescaled from 0 to 1 and the control group as the reference category.", fig.pos="H", fig.width=8, fig.height=6, message=FALSE, warning=FALSE, echo = FALSE, paged.print=FALSE}
# H1: misinformation effect
pdata_h1_belief_policy <- rbind(get_pdata(lm_moral_conviction, c(4, 7), "Moral conviction"),
                                get_pdata(lm_policy_opinion, c(4, 7), "Policy restrictions"),
                                get_pdata(lm_knowledge_safe, c(4, 7), "Belief GMO safety"))

pdata_h1_emotions <- rbind(get_pdata(lm_emotion_sadness, c(4, 7), "Sadness"),
                           get_pdata(lm_emotion_disgust, c(4, 7), "Disgust"),
                           get_pdata(lm_emotion_anxiety, c(4, 7), "Anxiety"),
                           get_pdata(lm_emotion_anger, c(4, 7), "Anger")
                           )

pdata_h1 <- rbind(cbind(pdata_h1_belief_policy, type = "Effect on belief and policy opinion"),
                  cbind(pdata_h1_emotions, type = "Effect on emotions"))

pdata_h1 <- pdata_h1 %>% 
  mutate(emotion = ifelse(str_detect(group, "low-misinfo"), 
                          "Low emotional misinformation", "High emotional misinformation"))

# actual plotting
ggplot(pdata_h1, aes(x = outcome, y = estimate, color = "1")) +
  geom_point(size = 1.5, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = lower, ymax = upper, linetype = emotion),
                size = 0.5,
                width = 0.25,
                position = position_dodge(width = 0.5)) +
  geom_hline(yintercept = 0, linetype = 3) +
  ylim(c(-0.2, 0.2)) +
  labs(x = "",
       y = "") +
  facet_wrap(fct_rev(emotion) ~ fct_rev(type), scales = "free") +
  coord_flip() +
  #scale_colour_manual(values = c("grey25", "grey50", "grey75")) +
  scale_color_viridis_d(begin = 0, end = 0.5, direction = 1) +
  scale_linetype_manual(values=c("solid", "twodash")) +
  scale_shape_manual(values = c(17, 16)) +
  theme_nice() +   
  theme(legend.title = element_blank(),
        legend.position = "bottom",
        legend.key.size = unit(1, 'cm'),
        strip.text.x = element_text(size = 8)
  ) +
  guides(color = "none",
         linetype = guide_legend(reverse = TRUE),
         shape = guide_legend(reverse = TRUE))

#sjPlot::save_plot("./figures/p_h1.png", width = 24, height = 16)

```

## Effectiveness of neutral correction {-}

```{r fig-2, fig.align="center", fig.cap="Coefficients (and 95 confidence intervals) from OLS-regressions with the outcome variable rescaled from 0 to 1 and the control group as the reference category.", fig.pos="H", fig.width=8, fig.height=6, message=FALSE, warning=FALSE, echo = FALSE, paged.print=FALSE}

# neutral correction for both low and high misinformation

pdata_neutral_correction_belief_opinion <- rbind(get_pdata(lm_moral_conviction, c(4, 5, 7, 8), "Moral conviction"),
                                           get_pdata(lm_policy_opinion, c(4, 5, 7, 8), "Policy restrictions"),
                                           get_pdata(lm_knowledge_safe, c(4, 5, 7, 8), "Belief GMO safety"))

pdata_neutral_correction_emotions <- rbind(
  get_pdata(lm_emotion_sadness, c(4, 5, 7, 8), "Sadness"),
  get_pdata(lm_emotion_disgust, c(4, 5, 7, 8), "Disgust"),
  get_pdata(lm_emotion_anxiety, c(4, 5, 7, 8), "Anxiety"),
  get_pdata(lm_emotion_anger, c(4, 5, 7, 8), "Anger")
  )

pdata_neutral_correction <- rbind(cbind(pdata_neutral_correction_belief_opinion, type = "Effect on belief and policy opinion"),
                            cbind(pdata_neutral_correction_emotions, type = "Effect on emotions"))

pdata_neutral_correction <- pdata_neutral_correction %>% 
  mutate(emotion = ifelse(str_detect(group, "low-misinfo"), "Low emotional misinformation", "High emotional misinformation"),
         correction = ifelse(str_detect(group, "correction"), "Neutral correction", "No correction"))

# actual plotting
ggplot(pdata_neutral_correction, aes(x = outcome, y = estimate, color = correction, shape = correction)) +
  geom_point(size = 1.5, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = lower, ymax = upper, linetype = emotion),
                size = 0.5,
                width = 0.5,
                position = position_dodge(width = 0.5)) +
  geom_hline(yintercept = 0, linetype = 3) +
  ylim(c(-0.2, 0.2)) +
  labs(x = "",
       y = "") +
  facet_wrap(fct_rev(emotion) ~ fct_rev(type), scales = "free") +
  coord_flip() +
  #scale_colour_manual(values = c("grey25", "grey50", "grey75")) +
  scale_color_viridis_d(begin = 0, end = 0.5, direction = -1) +
  scale_linetype_manual(values=c("solid", "twodash")) +
  scale_shape_manual(values = c(17, 16)) +
  theme_nice() +   
  theme(legend.title = element_blank(),
        legend.position = "bottom",
        legend.key.size = unit(1, 'cm'),
        strip.text.x = element_text(size = 8)
    ) +
  guides(color = guide_legend(reverse = TRUE),
         linetype = guide_legend(reverse = TRUE),
         shape = guide_legend(reverse = TRUE))

#sjPlot::save_plot("./figures/p_neutral_correction.png", width = 24, height = 16)
```

## Counter-frame {-}

```{r fig-3, fig.align="center", fig.cap="Coefficients (and 95 confidence intervals) from OLS-regressions with the outcome variable rescaled from 0 to 1 and the control group as the reference category.", fig.pos="H", fig.width=8, fig.height=6, message=FALSE, warning=FALSE, echo = FALSE, paged.print=FALSE}

# counter-frame correction for both low and high misinformation
pdata_counter_frame_belief_opinion <- rbind(
  get_pdata(lm_moral_conviction, c(4, 6, 7, 9), "Moral conviction"),
  get_pdata(lm_policy_opinion, c(4, 6, 7, 9), "Policy restrictions"),
  get_pdata(lm_knowledge_safe, c(4, 6, 7, 9), "Belief GMO safety")
  )

pdata_counter_frame_emotions <- rbind(
  get_pdata(lm_emotion_sadness, c(4, 6, 7, 9), "Sadness"),
  get_pdata(lm_emotion_disgust, c(4, 6, 7, 9), "Disgust"),
  get_pdata(lm_emotion_anxiety, c(4, 6, 7, 9), "Anxiety"),
  get_pdata(lm_emotion_anger, c(4, 6, 7, 9), "Anger")
  )

pdata_counter_frame <- rbind(cbind(pdata_counter_frame_belief_opinion, type = "Effect on belief and policy opinion"),
                                  cbind(pdata_counter_frame_emotions, type = "Effect on Emotions"))

pdata_counter_frame <- pdata_counter_frame %>% 
  mutate(emotion = ifelse(str_detect(group, "low-misinfo"), "Low emotional misinformation", "High emotional misinformation"),
         correction = ifelse(str_detect(group, "counter"), "Counter-frame correction", "No correction"))

# actual plot
ggplot(pdata_counter_frame, aes(x = outcome, y = estimate, color = correction, shape = correction)) +
  geom_point(size = 1.5, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = lower, ymax = upper, linetype = emotion),
                size = 0.5,
                width = 0.5,
                position = position_dodge(width = 0.5)) +
  geom_hline(yintercept = 0, linetype = 3) +
  ylim(c(-0.2, 0.2)) +
  labs(x = "",
       y = "") +
  facet_wrap(fct_rev(emotion) ~ fct_rev(type), scales = "free") +
  coord_flip() +
  #scale_colour_manual(values = c("grey25", "grey50", "grey75")) +
  scale_color_viridis_d(begin = 0, end = 0.75, direction = -1) +
  scale_linetype_manual(values=c("solid", "twodash")) +
  scale_shape_manual(values = c(15, 16)) +
  theme_nice() +   
  theme(legend.title = element_blank(),
        legend.position = "bottom",
        legend.key.size = unit(1, 'cm'),
        strip.text.x = element_text(size = 8)
  ) +
  guides(color = guide_legend(reverse = TRUE),
         linetype = guide_legend(reverse = TRUE),
         shape = guide_legend(reverse = TRUE))

#sjPlot::save_plot("./figures/p_counter_frame.png", width = 24, height = 16)

```

# Conclusion

Summary:

* Persuasive, emotionally framed misinformation can increase moralization of opinions about GM food.
* Effects persist with neutral corrective information, but counter-framing works.
* Overall small effects: Explained variance ($R^2$) of $7 \%$ for beliefs, $3.5 \%$ for opinions, $1 \%$ for moral conviction.

Limitations:

* is it really emotions or "just" the pictures?

Implications:

* When misinformation is emotionally charged, fact-checkers may have to think about counter-frames to eliminate the misinformation effect.
* a lack of acceptance of novel food technologies against scientific evidence can be seen as as a threat for the development of a more resilient food system @siegrist2020consumer.  genetic engineering provides avenues with large potential benefits @roberts2018nobel, @planck2019scientists, which may be impeded by public resistance possibly originating from misleading claims easily disseminated through online media.
* Similar effects possible on other issues where misinformation loads on strong emotions, e.g. health interventions such as vaccinations.

\newpage

# References

\linespread{1}
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\noindent

::: {#refs}

:::

\newpage
\pagenumbering{roman}

# Appendix {-}

## Information effect

```{r fig-4, fig.align="center", fig.cap="Coefficients (and 95 confidence intervals) from OLS-regressions with the outcome variable rescaled from 0 to 1 and the control group as the reference category.", fig.pos="H", fig.width=8, fig.height=6, message=FALSE, warning=FALSE, echo = FALSE, paged.print=FALSE}

# Appendix
# info only

pdata_info_effect <- rbind(
  get_pdata(lm_moral_conviction, c(2, 3), "Moral conviction"),
  get_pdata(lm_policy_opinion, c(2, 3), "Policy restrictions"),
  get_pdata(lm_knowledge_safe, c(2, 3), "Belief GMO safety"))

pdata_info_effect_emotions <- rbind(
  get_pdata(lm_emotion_sadness, c(2, 3), "Sadness"),
  get_pdata(lm_emotion_disgust, c(2, 3), "Disgust"),
  get_pdata(lm_emotion_anxiety, c(2, 3), "Anxiety"),
  get_pdata(lm_emotion_anger, c(2, 3), "Anger")
  )

pdata_info_effect <- rbind(cbind(pdata_info_effect, type = "Effect on belief and policy opinion"),
                           cbind(pdata_info_effect_emotions, type = "Effect on Emotions"))

pdata_info_effect <- pdata_info_effect %>% 
  mutate(group = fct_recode(group, 
                            "Neutral correction" = "neutral-correction",
                            "Counterframe" = "counter-frame"))

ggplot(pdata_info_effect, aes(x = outcome, y = estimate, color = fct_rev(group), shape = fct_rev(group))) +
  geom_point(size = 1.5, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                size = 0.5,
                width = 0.25,
                position = position_dodge(width = 0.5)) +
  geom_hline(yintercept = 0, linetype = 3) +
  ylim(c(-0.2, 0.2)) +
  labs(x = "",
       y = "") +
  facet_wrap(group~fct_rev(type), scales = "free_y") +
  coord_flip() +
  #scale_colour_manual(values = c("grey25", "grey50", "grey75")) +
  scale_color_viridis_d(begin = 0.25, end = 0.75, direction = -1) +
  scale_linetype_manual(values=c("dashed", "solid")) +
  scale_shape_manual(values = c(15, 17)) +
  theme_nice() +   
  theme(legend.title = element_blank(),
        legend.position = "bottom",
        legend.key.size = unit(1, 'cm'),
        strip.text.x = element_text(size = 8)
  ) +
  guides(color = guide_legend(reverse = TRUE),
         linetype = guide_legend(reverse = TRUE),
         shape = guide_legend(reverse = TRUE))

#sjPlot::save_plot("./figures/p_info_effect.png", width = 24, height = 9)

```

\newpage

## Regression tables

```{r reg-output-emotion, message=FALSE, warning=FALSE, echo = FALSE}

library(modelsummary)

cm <- c('(Intercept)'    = 'Control',
        'exp_grouplow-misinfo'    = 'Low emotion misinf.',
        'exp_grouphigh-misinfo' = 'High emotion misinf.',
        'exp_grouplow-misinfo-neutral-correction' = 'Neutral correction to low emotion misinf.',
        'exp_grouphigh-misinfo-neutral-correction' = 'Neutral correction to high emotion misinf.',
        'exp_grouplow-misinfo-counter-frame' = 'Counterframe low emotion misinf.',
        'exp_grouphigh-misinfo-counter-frame' = 'Counterframe to high emotion misinf.',
        'exp_groupneutral-correction'    = 'Neutral correction w/o misinf.',
        'exp_groupcounter-frame'    = 'Counterframe w/o misinf.')

models_emotions <- list("Anger" = lm_emotion_anger,
                        "Anxiety" =  lm_emotion_anxiety,
                        "Disgust" = lm_emotion_disgust,
                        "Sadness" = lm_emotion_sadness)

modelsummary(models_emotions,
             coef_map = cm,
             title = 'Regression table for emotions.', 
             stars = c('*' = .05, '**' = .01, '***' = 0.01),
             gof_omit = 'DF|Deviance|F|Log|AIC|BIC',
             output = 'markdown')

```
\newpage

```{r reg-output-belief-opinion, message=FALSE, warning=FALSE, echo = FALSE}

models <- list("Belief GMO safety" = lm_knowledge_safe,
               "Policy restrictions" =  lm_policy_opinion,
               "Moral conviction" = lm_moral_conviction)

modelsummary(models,
             coef_map = cm,
             title = 'Regression table for belief and policy opinion.', 
             stars = c('*' = .05, '**' = .01, '***' = 0.01),
             gof_omit = 'DF|Deviance|F|Log|AIC|BIC',
             output = 'markdown')

```

\newpage

## Treatment material

```{r screenshot-low-misinfo, fig.align = 'center', out.width = "100%", fig.cap = "Low emotion misinformation", echo = FALSE}
knitr::include_graphics("./figures/misinfo_low_emotion.png")
```

```{r screenshot-high-misinfo, fig.align = 'center', out.width = "100%", fig.cap = "High emotion misinformation", echo = FALSE}
knitr::include_graphics("./figures/misinfo_high_emotion.png")
```

```{r screenshot-neutral-correction, fig.align = 'center', out.width = "100%", fig.cap = "Neutral correction", echo = FALSE}
knitr::include_graphics("./figures/neutral_correction.png")
```

```{r screenshot-counterframe, fig.align = 'center', out.width = "100%", fig.cap = "Counterframe correction", echo = FALSE}
knitr::include_graphics("./figures/counter_frame.png")
```